---
title: "R Notebook"
output: html_notebook
---

## MySQL

## Question 1

### Connecting to MySQL server on Amazon AWS

```{r}
library(RMySQL)

db_user <- 'admin'
db_password <- 'khoury5200'
db_name <- 'practicum'
db_host <- 'parvpracticum.clwtsa2zfujt.us-east-2.rds.amazonaws.com'
db_port <- 3306

mydb <-  dbConnect(RMySQL::MySQL(), user = db_user, password = db_password,
                 dbname = db_name, host = db_host, port = db_port)
```

Dropping table if they already exist.

```{sql connection=mydb}
DROP TABLE IF EXISTS incidents, airports, conditions;
```

Create table incidents with unique constraints on origin and condition so they are unique as they are referenced by foreign key.

```{sql connection=mydb}
CREATE TABLE incidents (
  rid INT NOT NULL,
  `date` DATE NOT NULL,
  origin int NOT NULL,
  airline TEXT DEFAULT NULL,
  aircraft TEXT DEFAULT NULL,
  flightPhase TEXT DEFAULT NULL,
  altitude TEXT DEFAULT NULL,
  conditions varchar(200) NOT NULL,
  warning BOOLEAN DEFAULT NULL,
  PRIMARY KEY (rid),
  
  CONSTRAINT uq_orig UNIQUE (origin),
  CONSTRAINT uq_condition UNIQUE (conditions)
);
```

Creating table airports and also added foreign key aid which references to origin in incidents table.

```{sql connection=mydb}
CREATE TABLE airports (
  aid int NOT NULL,
  airportName varchar(100) NOT NULL, 
  airportCode varchar(5) DEFAULT NULL,
  state text,
  PRIMARY KEY (aid),
  CONSTRAINT aid FOREIGN KEY (aid) REFERENCES incidents(origin)
);
```

Created table conditions and added a foreign key which references to conditions in incidents table.

```{sql connection=mydb}
CREATE TABLE conditions (
  cid int NOT NULL primary key,
  conditions varchar(200) NOT NULL,
  explanantion text DEFAULT NULL,
  CONSTRAINT conditions FOREIGN KEY (conditions) REFERENCES incidents (conditions)
);
```

Reading the Bird Strikes CSV file.

```{r, warning=FALSE}
library(dplyr)
library(sqldf)

data.read <- read.csv(file = "BirdStrikesData-V2.csv")

data.read
```

Removing Military airlines as asked in the question using <> (not equal to) sign.

```{r}
detach("package:RMySQL", unload=TRUE)
data.read <- sqldf::sqldf("select * from `data.read` where airline <> 'MILITARY'")

data.read
```

Checking all the unique flight phases preset in the table.

```{r}
unique(data.read[c("flight_phase")])
```

Replacing the flight phases with the harmonizations given in the question. Also, replacing the warning value with True and False.

```{r}
data.read$flight_phase <- replace(data.read$flight_phase, 
                                  data.read$flight_phase == "Climb",
                                  "takeoff")

data.read$flight_phase <- replace(data.read$flight_phase, 
                                  data.read$flight_phase == "Take-off run",
                                  "takeoff")

data.read$flight_phase <- replace(data.read$flight_phase, 
                                  data.read$flight_phase == "Landing Roll",
                                  "landing")

data.read$flight_phase <- replace(data.read$flight_phase, 
                                  data.read$flight_phase == "Approach",
                                  "landing")

data.read$flight_phase <- replace(data.read$flight_phase, 
                                  data.read$flight_phase == "Descent",
                                  "landing")

data.read$flight_phase <- replace(data.read$flight_phase, 
                                  data.read$flight_phase == "",
                                  "unknown")

data.read$flight_phase <- replace(data.read$flight_phase, 
                                  data.read$flight_phase == "Taxi",
                                  "Parked")

data.read$pilot_warned_flag <- replace(data.read$pilot_warned_flag,
                                       data.read$pilot_warned_flag == "N",
                                       FALSE)

data.read$pilot_warned_flag <- replace(data.read$pilot_warned_flag,
                                       data.read$pilot_warned_flag == "Y",
                                       TRUE)

data.read
```

## Question 2

Extracting the colums needed for airports table with the help of sqldf package. Also added a synthetic primary key as aid.

```{r load airports}
library(sqldf)

data.airports <- sqldf::sqldf("SELECT DISTINCT airport AS airportName, 1 AS aid,
                               origin AS state FROM `data.read`")

n.airports <- nrow(data.airports)
data.airports[, 2] <- seq(1, n.airports)

data.airports <- data.airports %>%
  mutate(airportCode=NA,
         .before=state)

data.airports <- data.airports %>% select(aid, airportName, airportCode, state)

data.airports
```

Extracted the incidents table from the data with the help of sqldf package and also changed the datatype of date. Also, origin table is converted to aid.

```{r load incidents, warning=FALSE}
data.incidents <- sqldf::sqldf("SELECT rid, flight_date AS `date`,
                               airport AS origin, airline, aircraft, flight_phase AS flightPhase,
                               altitude_ft AS altitude, sky_conditions AS conditions, 
                               pilot_warned_flag AS warning FROM `data.read`")

data.incidents$date <- as.Date(data.incidents$date, "%m/%d/%Y")

ori <- sqldf::sqldf("SELECT a.aid, i.origin from `data.airports` as a 
                                      left join `data.incidents` as i
                                      on i.origin = a.airportName
                    GROUP BY aid")

for (i in 1:nrow(data.incidents)){
  data.incidents$origin[i] <- ori$aid[data.incidents$origin[i] == ori$origin]
}

data.incidents$origin <- as.integer(data.incidents$origin)

data.incidents
```

Extracted the conditions table from the csv file using sqldf package. Also, added the synthetic primary key.

```{r load conditions}
data.conditions <- sqldf::sqldf("SELECT DISTINCT sky_conditions as conditions, 1 as cid
                                FROM `data.read`")

n.conditions <- nrow(data.conditions)
data.conditions[, 2] <- seq(1, n.conditions)

data.conditions <- data.conditions %>%
  mutate(explanation=NA)

data.conditions <- data.conditions %>% select(cid, conditions, explanation)

data.conditions
```

## Question 3

### Writing the tables and displaying them

```{r}
library(RMySQL)
dbWriteTable(mydb, "conditions", data.conditions, overwrite = T, row.names = F)
```

```{sql connection=mydb}
SELECT * FROM conditions;
```

```{r}
library(RMySQL)
dbWriteTable(mydb, "airports", data.airports, overwrite = T, row.names = F)
```

```{sql connection=mydb}
SELECT * FROM airports;
```

```{r}
library(RMySQL)
dbWriteTable(mydb, "incidents", data.incidents, overwrite = T, row.names = F)
```

```{sql connection=mydb}
SELECT * FROM incidents;
```

## Question 4

Grouping airlines and counting the rid to get the number of incidents for each airline.

```{sql connection=mydb}
select airline, count(rid) as incident_count 
             from incidents
             group by airline
             order by incident_count desc
             limit 10;
```

## Question 5

Counting the rid to get the incident count for each flight phase and then dividing it by 4 tp get the average incidents for the flight phase. Displayng only the top result.

```{sql connection=mydb}
select flightPhase, count(rid)/4 as avg_impact
             from incidents
             group by flightPhase
             order by avg_impact desc
             limit 1
```

## Question 6

Extracting the Month of all years and counting them with the help of rid. Month august has the most number of incidents.

```{sql connection=mydb}
select MONTH(`date`) as Month, count(rid) as Incident_count
from incidents
group by Month
order by Incident_count desc
```

## Question 7

Extracting the year count for the year 2005 to 2011 and then created a column plot for display.

```{r}
library(tidyverse)

year_table <- dbGetQuery(mydb, 
"Select YEAR(`date`) as Year, count(rid) as incident_count
from incidents
group by Year
having Year >= 2005 and Year <= 2011
order by Year;")

ggplot(data = year_table) + 
  geom_col(mapping = aes(x = factor(Year), y = incident_count)) +
  labs(x = "Year", y = "Incident Count", 
       title = "Incident Count from (2005 - 2011)") +
  theme_minimal()
```

## Question 8

Displaying the airports table.

```{sql connection=mydb}
Select * from airports;
```

Creating the stored procedure which will delete the aid given by the user.

```{sql connection=mydb}
CREATE PROCEDURE `removeBirdStrike`  
(IN ID INT)
BEGIN
DELETE FROM airports WHERE aid = ID;
END
```

Setting ID 1 to be deleted.

```{sql connection=mydb}
SET @ID = 1
```

Calling the stored procedure.

```{sql connection=mydb}
CALL `removeBirdStrike`
(@ID)
```

Displaying the updated airports table.

```{sql connection=mydb}
Select * from airports;
```

Disconnecting Database

```{r}
dbDisconnect(mydb)
```